---
title: "day4_comp_morning"
author: "Joe DeCesaro"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```


```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)
half_mile <- vector(mode = "numeric", length = length(quarter_splits) - 1)

# Write a for loop that returns the total time for any two consecutive quarter miles

for (i in 1:(length(quarter_splits) - 1)) {
  half_mile[i] <- quarter_splits[i] + quarter_splits[i + 1]
}
half_mile
```

### Simple Functions Practice
Basic function structure practice
```{r}
birddog_sum <- function(bird, dog) {
  
  if (bird > 50) {
    stop("Birds can't be greater than 50!")
  }
  
  if (bird < 10) {
    message("Cool little birdie")
  }
  
  if (dog < 0) {
    warning("Can dogs be < 0 pounds?")
  }
  
  add_birddog <- bird + dog
  return(add_birddog)
}


birddog_sum(bird = 49, dog = 5)

birddog_sum(bird = 9, dog = -5)

```

```{r}
say_hello <- function(friend, weekday) {
  print(paste("Hello", friend,"have a great", weekday))
}

say_hello(friend = "Denise", weekday = "Friday")

```

Print vs return example
```{r}
find_multiple <- function(a, b) {
  my_multiplication <- a * b
  print(my_multiplication)
}

final_value <- find_multiple(a = 10, b = 6)

final_value
```

### Functions with conditions!
```{r}
animal_age <- function(animal, age) {
  
  if (!animal %in% c("dog", "goat")) {
    stop("Animal must be a dog or a goat.")
  }
  
  if (animal == "dog") {
    creature_age <- age * 7
  }
  
  else if (animal == "goat") {
    creature_age <- age * 4.7
  }
  
  return(creature_age)
}

# animal_age(animal = "cat", age = 10)
# Above example breaks so it would not knit
```

Common misuse of the `%in%` operator
```{r}
adelie_gentoo <- penguins %>% 
  filter(species %in% c("Adelie", "Gentoo"))

adelie_gentoo_wrong <- penguins %>% 
  filter(species == c("Adelie", "Gentoo"))
# By writing the statement as above it is alternating between the two values in the vector until it gets to the end of the vector.
```

### Functions with for loops!

```{r}
df_means <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    print(paste("The mean value of", colnames(df[i]), "is", round(col_mean,2)))
  }
}

df_means(df = mtcars)

# When building functions test it on the simplest of data sets first

# Make a subset of only structural size variables from the penguins dataset for Chinstrap

chinstrap_sizes <- penguins %>% 
  filter(species %in% c("Chinstrap")) %>% 
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)

df_means(df = chinstrap_sizes)
  
```

### Condition to check the class of an input
```{r}
lunchtime <- function(tacos, price_per) {
  
  if (is.numeric(tacos) == FALSE) {
    stop("You must input a number of tacos.")
    
  }
  tot_price <- tacos * price_per
  return(tot_price)
}

lunchtime(tacos = 4, price_per = 2.50)
```

### Example where we use values from a data fram to calculate something based on user input
```{r}
# Calculate the groundwater transport rate

gw_site <- function(site) {
  
  gw_slope_depth <- data.frame(sitename = c("mountain", "prairie", "desert", "beach"), 
                                 depth = c(32, 41, 63, 2), 
                                 slope = c(11.2, 0.4, 0.8, 2.6))
  # this is another way to do the below: site_select <- dplyr::filter(gw_slope_depth, sitename == site)
  site_select <- dplyr::filter(gw_slope_depth, sitename == site)
  
  transport_rate <- 1.4 * site_select$slope + 3.6 * site_select$depth
  
  return(transport_rate)
}

gw_site(site = "beach")
gw_site(site = "mountain")

```

Below is the same example as above except using the switch() function
```{r}
gw_rate_switch <- function(site) {
  
  gw_slope_depth <- switch(site, 
                           "mountain" = c(32, 11.2),
                           "prairie" = c(41, 0.4),
                           "desert" = c(63, 0.8),
                           "beach" = c(2, 2.6))
  transport_rate <- 1.4 * gw_slope_depth[2] + 3.6 * gw_slope_depth[1]
  
  return(transport_rate)
  
}

gw_rate_switch(site = "mountain")
gw_rate_switch(site = "beach")
```

### Now let's do an example with the logistic growth function
```{r}
logistic_growth <- function(K, N0, r, time) {
 Nt <- K / (1 + ((K - N0) / N0) *exp(-r * time))
 print(Nt)
}

logistic_growth(K = 5000, N0 = 200, r = 0.43, time = 6)

time_vec <- seq(from = 0, to = 35, by = 0.1)
bacterial_growth <- logistic_growth(K = 5000, N0 = 200, r = 0.43, time = time_vec)

bacterial_df <- data.frame(time_vec, bacterial_growth)

ggplot(data = bacterial_df, aes(x = time_vec, y = bacterial_growth)) +
  geom_point()

```
Create a new sequence of R values that are a range
```{r}
r_seq <- seq(from = 0.20, to = 0.40, by = 0.01)
time_vec <- seq(from = 0, to = 35, by = 0.1)

out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

# N0 = 100, K = 6000

for (i in seq_along(r_seq)) {
  for (j in seq_along(time_vec)) {
    pop <- logistic_growth(K = 6000, 
                           N0 = 100, 
                           r = r_seq[i], 
                           time = time_vec[j])
    out_matrix[j,i] <- pop
  }
}

# Wrangling 
out_df <- data.frame(time = time_vec, out_matrix)

# Update column names
colnames(out_df) <- c("time", paste0("gr_", r_seq))

# make a long form (tidy version) of this data

out_df_long <- out_df %>% 
  pivot_longer(-time, names_to = "growth_rates", values_to = "population")

# Make in ggplot a line graph (geom_line()) for population (y-axis) over time (x-axis) with color depending on the growth rate

ggplot(data = out_df_long, aes(x = time, y = population)) +
  geom_line(aes(color = growth_rates))

```

